\section{Problem Setup}
\label{sec:setup}

% MANUSCRIPT-NODE: SETUP-P01
Let $\nu$ denote the empirical update distribution induced by offline data, replay, or stale trajectories. We study the signed actor field
\begin{equation}
\mathbf F(\theta)=\mathbb E_{(s,a)\sim\nu}\!\left[\widehat A(s,a)\nabla_\theta\log\pi_\theta(a\mid s)\right].
\label{eq:signed-field}
\end{equation}
Writing $A^+=\max(\widehat A,0)$ and $A^-=\max(-\widehat A,0)$ gives $\mathbf F=\mathbf F^+-\mathbf F^-$. For a negative sample,
\begin{equation}
\|g_i^-\|=A_i^-\,\|\nabla_\theta\log\pi_\theta(a_i\mid s_i)\|.
\label{eq:factorization}
\end{equation}
The first factor is sample badness or severity; the second is learner-relative policy geometry. Sample count, directional coherence, and the network Jacobian determine how individual terms aggregate. Equation~\ref{eq:factorization} motivates matching badness while changing only distance or rarity.
% END-MANUSCRIPT-NODE: SETUP-P01

% MANUSCRIPT-NODE: SETUP-P02
We use $r_\theta(s,a)$ as a family-indexed remoteness coordinate. For a Gaussian policy, $r_\theta$ is a registered standardized or Mahalanobis distance (or an equivalent calibrated negative log-density). For a categorical policy, it is action surprisal $-\log\pi_\theta(a\mid s)$; for sequence policies, the registered analogue is normalized token- or completion-level negative log-likelihood. The notation is shared, but the score-growth law is not assumed to be identical across policy families. Crucially, far field is dynamic: a fixed historical action can become remote as the current learner changes.
% END-MANUSCRIPT-NODE: SETUP-P02
