\section{Distributionally Robust Policy Optimization}
\label{sec:method}

% MANUSCRIPT-NODE: METHOD-P01
DRPO applies a policy-relative envelope to the negative component of the signed actor field. Let $r_\theta(s,a)\ge 0$ denote the registered remoteness measure for the policy family and experiment. We define
\begin{equation}
w^-_\lambda(s,a)=\exp[-\lambda r_\theta(s,a)],\qquad \lambda\ge 0,
\label{eq:exp-weight}
\end{equation}
and optimize
\begin{equation}
\mathbf F_{\mathrm{DRPO}}(\theta)=\mathbb E_\nu\!\left[A^+\nabla\log\pi_\theta-A^-w^-_\lambda\nabla\log\pi_\theta\right].
\label{eq:drpo-field}
\end{equation}
Positive updates are unchanged, nearby negative actions retain substantial influence, and increasingly remote negative actions are attenuated. Sample quality and learner-relative remoteness remain distinct axes: the matched experiments vary remoteness without changing badness, while Appendix~\ref{app:optimistic-dro} states the separate Optimistic-DRO quality-selection result. The main DRPO update does not equate the two controls.
% END-MANUSCRIPT-NODE: METHOD-P01

% MANUSCRIPT-NODE: METHOD-P02
In the exponential-family analysis, the uncontrolled negative moment is
\begin{equation}
q\mathbf m_-=\mathbb E[A^-T(a)].
\end{equation}
DRPO replaces it by
\begin{equation}
q_\lambda\mathbf m_{-,\lambda}=\mathbb E[A^-e^{-\lambda r_\theta(s,a)}T(a)],
\label{eq:weighted-negative-moment}
\end{equation}
which induces the controlled target
\begin{equation}
\mathbf m_\lambda^\star=\frac{p\mathbf m_+-q_\lambda\mathbf m_{-,\lambda}}{p-q_\lambda}.
\end{equation}
The method therefore modifies the same term that first creates stable extrapolation and later drives the signed target toward the boundary. Small $r_\theta$ preserves substantial negative influence; large $r_\theta$ is selectively attenuated.
% END-MANUSCRIPT-NODE: METHOD-P02

% MANUSCRIPT-NODE: METHOD-P03
\begin{proposition}[Vanishing weighted far-field gradient]
\label{prop:vanishing}
Suppose the policy score satisfies $\|\nabla_\theta\log\pi_\theta(a\mid s)\|\le C(1+r_\theta(s,a))^k$ for finite $C,k$. For any $\lambda>0$,
\begin{equation}
e^{-\lambda r_\theta(s,a)}\|\nabla_\theta\log\pi_\theta(a\mid s)\|\rightarrow 0
\quad\text{as }r_\theta(s,a)\rightarrow\infty.
\end{equation}
\end{proposition}
The exponential form is therefore a gradient-tail envelope derived from the far-field score-growth bound. Directional utility is measured separately in the controlled experiments.
% END-MANUSCRIPT-NODE: METHOD-P03

% MANUSCRIPT-NODE: METHOD-P04
We compare DRPO against an uncontrolled signed baseline, Positive-only learning, non-selective global scaling, registered linear or reciprocal tapers, and a hard threshold on the same remoteness coordinate. A quality-based hard filter is reported separately because it controls a different axis. Where the protocol calls for budget matching, methods receive the same pre-optimizer raw negative-gradient norm and are evaluated with paired seeds. No method ranking is assumed in advance; best and terminal checkpoints, local-negative retention, far-field retention, task performance, boundary events, and numerical failures are all reported.
% END-MANUSCRIPT-NODE: METHOD-P04
